---
title: "Econ294A - Final Exam"
author: "Feng Hong"
date: "March 16, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
library(Rcpp)
library(ggplot2)
library(gridExtra)
library(dplyr)
library(class)
library(nycflights13)  #for data
library("RSQLite") #for sqllite
```

## Extract raw data from sqlite

```{r raw, message=FALSE}
nycflights13_sqlite()

flights_sqlite <- tbl(nycflights13_sqlite(), "flights")
airlines_sqlite <- tbl(nycflights13_sqlite(), "airlines")
airports_sqlite <- tbl(nycflights13_sqlite(), "airports")
planes_sqlite <- tbl(nycflights13_sqlite(), "planes")
weather_sqlite <- tbl(nycflights13_sqlite(), "weather")
```

## join flights and planes data

```{r flights_planes_join, message=FALSE}

inner_flights_planes <- inner_join(flights, planes, by = "tailnum") %>% tbl_df 
colnames(inner_flights_planes)[1] <- "flight_year"
colnames(inner_flights_planes)[15] <- "dep_hour"
colnames(inner_flights_planes)[17] <- "plane_year"
#names(inner_flights_planes)
```

## create the date index

```{r data_index_creation}
inner_flights_planes <- inner_flights_planes %>% 
  mutate(
    date = paste(flight_year, month, day, sep = "-"), 
    date = as.Date(date, format = "%Y-%m-%d"), # create date to merge with weather
    cancelled = ifelse(is.na(arr_time), 1, 0)  # question requires this
  ) 

```

## select columns needed from the inner_flights_planes dataset

```{r select_columns}
flights_planes <- inner_flights_planes %>%
  dplyr::select(
    cancelled, date, month, day, dep_hour,
    dep_time, dep_delay, arr_time, arr_delay,
    carrier, flight, origin, dest, air_time, distance, 
    plane_year, manufacturer, seats)
```

- change character variable to factor variable
- change integer variable to factor variable

``` {r type_conversion}
flights_planes$carrier <- as.factor(flights_planes$carrier)
flights_planes$origin <- as.factor(flights_planes$origin)
flights_planes$dest <- as.factor(flights_planes$dest)
flights_planes$manufacturer <- as.factor(flights_planes$manufacturer)
flights_planes$month <- as.factor(flights_planes$month)
flights_planes$flight <- as.factor(flights_planes$flight)
```






##refine weather data

```{r weather1}
weather <- weather_sqlite %>%  
  collect() %>% 
  mutate(
    date = paste(year, month, day, sep = "-"),
    date = as.Date(date, format = "%Y-%m-%d"),
    weekday = weekdays(date),
    weekday = as.factor(weekday) # add the weekday variable
  ) 
```

```{r weather2}
weather_mean <- weather %>% group_by(date) %>%
  summarise(
    weekday = first(weekday),
    mean_temp = mean(temp),
    mean_dewp = mean(dewp),
    mean_humid = mean(humid),
    # mean_wind_dir = mean(wind_dir), wind direction has too many NA's.
    # mean_wind_speed = mean(wind_speed),
    # mean_wind_gust = mean(wind_gust), excluded because their effects depending on direction.
    mean_precip = mean(precip),
    # mean_pressure = mean(pressure), pressure has too many NA's.
    mean_visib = mean(visib)
  )
```



## identify the highly correlated data

```{r weather3}
corr_weather_mean <- cor(na.omit(weather_mean[,3:7]))
print(corr_weather_mean)
```

## delete columns that are highly correlated from weather_mean

```{r delete_colum}
weather2 <- weather_mean %>%
  dplyr::select(date, weekday, mean_temp, mean_precip, mean_visib)
```

##join flights, planes and weather data 

```{r join_all_table} 
final_data <- inner_join(flights_planes, weather2, by = "date")
length(final_data[final_data$cancelled == 1])
#names(final_data)
```


```
Now I run two regression model. One is a OLS model for predicting departure delay. 
The other is a GLS model (logit model) for predicting cancellation.
```


## **OLS** for *dep_delay
```{r OLS}
model.delay <- lm(dep_delay ~ month + as.factor(weekday) + carrier + origin +
                plane_year + seats + mean_precip + mean_visib,
               data = final_data)
summary(model.delay)
```

## **GLS** for *cancel*

```{r GLS}
model.cancel <- glm(cancelled ~ month + as.factor(weekday) + carrier + origin +
                     seats +  mean_visib,
                   data = final_data, family=binomial(link="logit"))
summary(model.cancel)
```

```
The regressors chosen are largely significant as we can see from the results.
Two plots are generated to illustrate some relationship between cancellation and other
conditions as well as departure delay and other conditions.
```

## (a) weather 

```{r plotting}
plot_weather <- ggplot(data = final_data, aes(mean_precip, mean_visib)) 
plot_weather + geom_point( aes(color = as.factor(cancelled)),size = 1) +
  xlab("Precipitation") + ylab("Visibility")
```


## (b) day of week and time of year

```{r day_of_week}
month_weekday <- final_data %>%
  group_by(month, weekday) %>%
  summarise(mean.dep_delay = mean(dep_delay, na.rm = T)) 

plot_time <- ggplot(month_weekday, aes(x = month, y= mean.dep_delay))
plot_time + geom_point(aes(color = weekday), size = 3) +
  xlab("Month") + ylab("Departure Delay")
```